Comparing the Direction and Degree of Selection on Mutations Between SARS-CoV-2 Phylogenetic Clades

### 2022-10-26 notebooks/expected_and_actual_counts_ratio.Rmd
Add an additional method for calculating p-values to compare speed

---
title: "Comparing the Direction and Degree of Selection on Mutations Between SARS-CoV-2 Phylogenetic Clades"
output: html_document
date: Sys.time()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

# Define a minimum threshold for the number of expected counts. All mutations below this threshold will be removed.
min_expected <- 10

@@

For the analyses conducted in this notebook, subset the data by "all" to include data collected from all countries. Additionally, remove low-quality mutations or branches with abnormally high mutation rates. Finally, create a column with mutation type (there are 12 possible mutations).

```{r filter_all_data}
# Include data from all countries AND exclude branches with very high mutation rates and questionable mutations.
filtered_SARS_mutation <- subset(SARS_mutation, subset=="all" & exclude=="FALSE")

@@

```{r pseudocount}
# Calculate the ratio of actual counts (i.e., observed) to expected counts (i.e., expected) with pseudocount of 0.5.
filtered_SARS_mutation$observed_expected <- (filtered_SARS_mutation$actual_count + pseudo)/(filtered_SARS_mutation$expected_count + pseudo)
```

@@

Describe the entire data set.

@@

```{r nucleotide_site, echo=FALSE}
ggplot(filtered_SARS_mutation, aes(color=mutation_type, y=actual_count/expected_count, x=nt_site))+
  geom_point()+
  theme_bw()+
  labs(x="", y="")
```

@@

@@

```{r contingency_analysis}
# Save list of p-values for every comparison. NOTE: Expected counts are not integers and are therefore rounded for the Fisher's Exact test.
'p_values <- data.frame(p_value=apply(filtered_comparisons, 1, FUN=function(c) {fisher.test(fisher_data[c(which(fisher_data$nt_mutation_clade==c[1]), which(fisher_data$nt_mutation_clade==c[2])), c("observed","expected")])$p.value}))'

>>>>

---
title: "Comparing the Direction and Degree of Selection on Mutations Between SARS-CoV-2 Phylogenetic Clades"
output: html_document
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

@@

# Define a minimum threshold for the number of expected counts. Mutations below this threshold will be removed.
min_expected <- 10

@@

For the analyses conducted in this notebook, subset the data by the country(countries) the sequence data came from. Additionally, remove low-quality mutations or branches with abnormally high mutation rates. Finally, create a column with mutation type (there are 12 possible mutations).

```{r filter_all_data}
# Include data from certain countries AND exclude branches with very high mutation rates and questionable mutations.
filtered_SARS_mutation <- subset(SARS_mutation, subset==country & exclude=="FALSE" & expected_count>=min_expected)

@@

```{r pseudocount}
# Calculate the ratio of actual counts (i.e., observed) to expected counts (i.e., expected) with a pseudocount.
filtered_SARS_mutation$observed_expected <- (filtered_SARS_mutation$actual_count + pseudo)/(filtered_SARS_mutation$expected_count + pseudo)
```

@@

Describe the filtered data set.

@@

```{r nucleotide_site, echo=FALSE}
ggplot(filtered_SARS_mutation, aes(color=mutation_type, y=log(actual_count/expected_count), x=nt_site))+
  geom_point()+
  theme_bw()+
  labs(x="", y="")
```

@@

```{r temp}
filtered_comparisons$clade1_observed <- fisher_data[which(fisher_data$nt_mutation_clade %in% filtered_comparisons$clade1),"observed"]
filtered_comparisons$clade1_expected <- fisher_data[which(fisher_data$nt_mutation_clade %in% filtered_comparisons$clade1),"expected"]
filtered_comparisons$clade2_observed <- fisher_data[which(fisher_data$nt_mutation_clade %in% filtered_comparisons$clade2),"observed"]
filtered_comparisons$clade2_expected <- fisher_data[which(fisher_data$nt_mutation_clade %in% filtered_comparisons$clade2),"expected"]

# alternate p-value calculations between sequential rows
p_values_alt <- data.frame(p_value=apply(filtered_comparisons, 1, FUN=function(c) {fisher.test(data.frame(clade1=c(as.numeric(c[3]),as.numeric(c[4])),clade2=c(as.numeric(c[5]),as.numeric(c[6]))))$p.value}))
```

@@

```{r contingency_analysis}
# Save list of p-values for every comparison. NOTE: Expected counts are not integers and are therefore rounded for the Fisher's Exact test. Mac Desktop Local: "334.273  41.641 375.678"
p_values <- data.frame(p_value=apply(filtered_comparisons, 1, FUN=function(c) {fisher.test(fisher_data[c(which(fisher_data$nt_mutation_clade==c[1]), which(fisher_data$nt_mutation_clade==c[2])), c("observed","expected")])$p.value}))
